\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-09-19 introduction.tex
Starting to edit

\documentclass[letterpaper, 11pt]{IEEEtran}      % Use this line for a4 paper

%LatexiDiff ignore tikz:
% $: latexdiff -c "PICTUREENV=(?:picture|tikzpicture|DIFnomarkup)[\w\d*@]*" old.tex new.tex > diff.tex

%\IEEEoverridecommandlockouts                              %

%\usepackage{mathcomSTEP}

%\overrideIEEEmargins 
%\IEEEoverridecommandlockouts                              % This command is only needed if 
% you want to use the \thanks command

%\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\onecolumn

\usepackage{mathptmx} 
\usepackage{times} 
\usepackage{amsmath} 
\usepackage{amsbsy} 
\usepackage{amssymb}

%\usepackage{newtxtext, newtxmath}
\usepackage{mathrsfs}
\usepackage{comment}
\usepackage[export]{adjustbox}
\usepackage{tikz}
\usetikzlibrary{external,positioning,decorations.pathreplacing,shapes,arrows,patterns}

\usepackage[hidelinks]{hyperref}
%\tikzexternalize[mode=list and make]
\usepackage{algorithmicx}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}
\usepackage[latin1]{inputenc}
\usetikzlibrary{arrows,shapes}
\usepackage{xifthen}
\usepackage{epic}
%\usepackage{caption}
\usepackage{epstopdf}

\newtheorem{thm}{\bf{Theorem}}
\newtheorem{cor}[thm]{\bf {Corollary}}
\newtheorem{lem}[thm]{\bf {Lemma}}
\newtheorem{prop}[thm]{\bf {Proposition}}
\newtheorem{example}{\bf {Example}}
\newtheorem{definition}{\bf {Definition}}
\newtheorem{rem}{\bf {Remark}}
\newtheorem{assump}{\bf {Assumptions}}

\newcommand{\ex}[1]{\ensuremath{\mathbb{E}\left[ #1\right]}}

\newcommand{\Var}[1]{\ensuremath{\mathrm{Var} \left[ #1\right]}}

\newcommand{\unif}{\mathsf{unif}}
\newcommand{\card}{\mathrm{card}}
\newcommand{\ARE}{\mathsf{ARE}}
\newcommand{\supp}{\mathrm{supp} }
\renewcommand\vec[1]{\ensuremath\boldsymbol{#1}}
\newenvironment{proof}{\paragraph*{Proof}}{\hfill$\square$ \newline}
\newcommand{\sgn}{\mathrm{sgn} }
\newcommand{\Prob}{\mathbb{P} }
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand*{\QEDA}{\hfill\ensuremath{\square}}
\newcommand{\CEO}{\mathsf{CEO}}
\newcommand{\Ncal}{\mathcal{N}}

\newcommand{\enc}{\mathsf{enc}}
\newcommand{\est}{\mathsf{est}}

\tikzstyle{int}=[draw, fill=blue!10, minimum height = 1cm, minimum width=1.5cm,thick ]
\tikzstyle{sint}=[draw, fill=blue!10, minimum height = 0.5cm, minimum width=0.8cm,thick ]
\tikzstyle{sum}=[circle, fill=blue!10, draw=black,line width=1pt,minimum size = 0.5cm, thick ]
\tikzstyle{ssum}=[circle, fill=blue!10,draw=black,line width=1pt,minimum size = 0.1cm]
\tikzstyle{int1}=[draw, fill=blue!10, minimum height = 0.5cm, minimum width=0.7cm,thick ]
\tikzstyle{enc}=[draw, fill=blue!10, minimum height = 2.7cm, minimum width=1cm,thick ]
\tikzstyle{int}=[draw, fill=blue!10, minimum height = 1cm, minimum width=1.5cm,thick ]

\author{
    \IEEEauthorblockN{
Alon Kipnis\IEEEauthorrefmark{1} and   
    John C. Duchi\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}  \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Stanford University, Department of Statistics \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Stanford University, Department of Electrical Engineering. }
}

\title{\LARGE \bf Mean Estimation from One-Bit Measurements}

\begin{document}
\graphicspath{{./Figs/}}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a symmetric log-concave distribution under the following constraint: only a single bit per sample from this distribution is available to the estimator. We study the mean squared error risk in this estimation as a function of the number of samples, and hence the number of bits, from this distribution. 
%
Under an adaptive setting in which each bit is a function of the current sample and previously recorded bits, we show that the optimal relative efficiency compared to the sample mean is the efficiency of the median. %For example, in estimating the mean of a normal distribution, a constraint of one bit per sample incurs a penalty of $\pi/2$ in sample size compared to the unconstrained case. 
%
We also consider a distributed setting where each bit is only a function of a single sample. We show that the optimal efficiency of the adaptive setting can be attained by splitting the sample into two parts and allowing only a single adaptation. Otherwise, we show that no one sample estimation procedure can attain the optimal efficiency uniformly over all points in the parameter space.
%
%Finally, we analyze a distributed setting where the bits are obtained by comparing each sample against a prescribed threshold. Consequently, we consider the threshold density that minimizes the maximal MSE. 
%
Our results indicate that estimating the mean from one-bit measurements is equivalent to estimating the sample median from these measurements. In the adaptive case, this estimate can be done with vanishing error for any point in the parameter space. In the distributed case, this estimate can be done with vanishing error only for a finite number of possible values for the unknown mean. 
\end{abstract}

%{\color{red}  See LASSO results in http://arxiv.org/abs/1506.02181v1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}
Estimating parameters from data collected and processed by multiple units may be limited due to communication constraints between these units. 
%
For example, this scenario arises in sensor arrays where information is collected at multiple physical locations and transmitted to a central estimation unit. In these situations, the ability to estimate a particular parameter from the data is dictated not only by the quality of observations and their number but also by the available bandwidth for communicating between the sensors and the central estimator. The question that we ask is to what extent a parametric estimation task is affected by this constraint on communication, and what are the fundamental performance limits in estimating a parameter subject to such restriction. \par
%%
\begin{figure*}
\begin{center}
\begin{tikzpicture}[node distance=2cm,auto]

  \node at (0,0) (source) {$X_1$} ;
 \node[below of = source, node distance = 1.5cm] (source2) {$X_2$};
\node[below of = source2, node distance = 2.1cm] (source3) {$X_n$};

\node[int1, right of = source2, node distance = 1.2cm] (enc2) {$\enc$};

\draw[->,line width = 2pt] (source2) -- (enc2); 
\draw[->,line width = 2pt] (source) -| (enc2); 
\draw[->,line width = 2pt] (source3) -| (enc2);

\node[below of = source2, node distance = 0.5cm] {$\vdots$};

\draw[->,line width = 2pt] (source2) -- (enc2); 
\node[int1, right of = enc2, node distance = 2.7cm ] (est) {$\est$};

\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {\small $B_1,\ldots,B_n$} (est);

\node[right of = est, node distance = 1.2cm] (dest) {\small ${\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=1pt] (est) -- (dest);
\node at (2,-4.5) {(i) Centralized};
\end{tikzpicture}
%
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {\small $X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {$\enc$};  
\draw[->,line width = 2pt] (source) -- (enc1);

 \node[below of = source, node distance = 1.5cm] (source2) {\small $X_2$};
\node[int1, right of = source2, node distance = 1.2cm] (enc2) {$\enc$};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source2, node distance = 2.1cm] (source3) {\small $X_n$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {$\enc$};  
\draw[->,line width = 2pt] (source3) -- (enc3);

\node[below of = source2, node distance = 0.5cm] {$\vdots$};
%\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim P_X$};

\node[int1, right of = enc2, node distance = 3cm ] (est) {$\est$};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1cm] (mes1) {$B_1$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$B_2$} (est);

\draw[->] (enc1)+(0.7,0) -- +(0.7,-0.7) -| (enc2);

%\draw[dashed,->] (enc2)+(0.7,0) -- +(0.7,-0.7) -| +(0,-1.2);
\draw[->,line width = 2pt] (enc3) -| (est);

%\node[below of = mes2, node distance = 1.5cm] (mes3) {$B_1,\ldots,B_{n-1} $};

\draw[->,line width = 2pt] (enc3) -| node[above, xshift = -1cm]  {$B_n$} (est);

%\draw[->] (mes3.west) -- +(0,-0.5) -| (enc3);

\node[below right = 2.3cm and 0.7cm of 
enc1.center, yshift = 0.cm] (end) {$B_1,\ldots,B_{n-1}$};

\draw[->] (enc1)+(0.7,0) -- (end.west) -| (enc3);

\node[right of = est, node distance = 1.2cm] (dest) {${\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=1pt] (est) -- (dest);
\node at (2,-4.5) {(ii) Adaptive};
\end{tikzpicture}
%
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {$X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {$\enc$};  
\draw[->,line width = 2pt] (source) -- (enc1);

 \node[below of = source, node distance = 1.5cm] (source2) {$X_2$};
\node[int1, right of = source2, node distance = 1.2cm] (enc2) {$\enc$};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source2, node distance = 2.1cm] (source3) {$X_n$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {$\enc$};  
\draw[->,line width = 2pt] (source3) -- (enc3);

%\node[above of = source3, node distance = 1cm] (dist) {$X_i \sim {\Ncal} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};

\node[int1, right of = enc2, node distance = 2.1cm ] (est) {$\est$};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1cm] (mes1) {$B_1$} (est);

\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$B_2$} (est);

\draw[->,line width = 2pt] (enc3) -| (est);

\draw[->,line width = 2pt] (enc3) -| node[above, xshift = -1cm]  {$B_n$} (est);

\node[right of = est, node distance = 1.2cm] (dest) {${\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=1pt] (est) -- (dest);
\node at (2,-4.5) {(iii) Distributed};
\end{tikzpicture}
\end{center}
\caption{\label{fig:setup} 
Three encoding settings: (i) Centralized  -- an encoder sends $n$ bits after observing $n$ samples. (ii) Adaptive (sequential) --  the $i$th encoder sends the bit $B_i$ depending on its private sample $X_i$ and previous bits $B_1,\ldots,B_{i-1}$. (iii) Distributed  -- each encoder send the bit $B_i$ based on its private sample $X_i$ only.}
\end{figure*}
%%

This paper answers this question in a particular setting: the estimation of the mean $\theta$ of a symmetric log-concave distribution with finite variance, under the constraint that only a single bit can be communicated on each sample from this distribution. As it turns out, the ability to share information before committing on each one-bit message dramatically affects the performance in estimating $\theta$. We, therefore, distinguish among the three settings illustrated in Figure~\ref{fig:setup}
\begin{itemize}
 \item[(i)\,\,] \emph{Centralized} encoding: all $n$ encoders confer and produce a single message consists of $n$ bits.
 \item[(ii)\,] \emph{Adaptive} or \emph{sequential} encoding: The $n$th encoder observes the $n$th sample and the $n-1$ previous bits.
 \item[(iii)] \emph{Distributed} encoding: The $n$th message is only a function of the $n$th sample.
 \end{itemize}
Evidently, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). Below are three application examples for each of the settings (i)-(iii) above, respectively:
\begin{itemize}
\item {\bf Signal acquisition:} A quantity is measured $n$ times at different instances. The results are averaged in order to reduce measurement noise and the averaged result is then stored or communicated using $n$ bits. 
\item {\bf Analog-to-digital conversion:} A sigma-delta modulator (SDM) converts an analog signal into a sequence of bits by sampling the signal at a very high rate and then using one-bit threshold detector combined with a feedback loop to update an accumulated error state \cite{1092194}. Therefore, the expected error in tracking an analog signal using an SDM falls under our setting (ii) when we assume that the signal at the input to the modulator is a constant (direct current) corrupted by, say, thermal noise \cite{53738}. Since the sampling rates in SDM are usually many times more than the bandwidth of its input, analyzing SDM under a constant input provides meaningful lower bound even for non-constant signals.
\item {\bf Privacy:} A business entity is interested in estimating the average income of its clients. In order to keep this information as confidential as possible, each client independently provides an answer to a yes/no question related to its income.
\end{itemize}
%
We measure the performance in estimating $\theta$ by the quadratic risk. For an estimator with finite quadratic risk $R_n$, we are interested in particular in the limit
\begin{equation}
\label{eq:ARE_def}
\lim_{n\to \infty} \frac{\sigma^2/n}{R_n},
\end{equation}
describing the \emph{asymptotic relative efficiency} (ARE) of this estimator compared to an estimator whose variance decreases as $\sigma^2/n+o(n^{-1})$. Estimators of this form include the empirical mean of the samples, and, under some conditions, the optimal Bayes estimator. \par
%
In addition to the examples above, the excess risk in estimating a fixed parameter due to a one-bit per sample constraint is useful in bounding from below the excess risk in estimating signals from thier noisy one-bit quanitzed noisy measurements. Specifically, the excess risk or ARE we derive serve as the most optimistic estimate for the risk in estimating under one-bit per measurement constraint signals changing over time or space. Such estimation settings are considered in \cite{baraniuk2017exponential, jacques2013robust, plan2013one, li2017channel, choi2016near}. \par

@@

\section{Problem Formulation \label{sec:problem}}

Let $f(x)$ be a symmetric and log-concave density function with a finite second moment $\sigma^2$. For $\theta \in \Theta$, denote by $P_X$ the probability distribution with density $f\left( x-\theta \right)$. Therefore, $P_X$ is an absolutely continuous log-concave distribution with mean $\theta$ and variance $\sigma^2$. Symmetry and log-concavity of $f(x)$ imply that $P_X$ is strongly unimodal (has a unique local maxima) with its mode at $x =\theta$ \cite{ibragimov1956composition}. We further assume that the \emph{parameter space} $\Theta$ is a closed interval of the real line. 
\par
In some situations it is useful to assume that $\theta$ is drawn once from the prior distribution $\pi$ on $\Theta$. In this case we assume that $\pi$ is an absolutely continuous distribution, and denote its density by $\pi(\theta)$, i.e., $\pi(d\theta) = \pi(\theta)d\theta$.  
\par
The random variables $X_1,\ldots,X_n$ represent $n$ independent samples from $P_X$. 
We are interested in estimating $\theta$ from a set of $n$ binary messages $B_1,\ldots,B_n$, obtained from $X_1,\ldots,X_n$ under three possible scenarios illustrated in Figure~\ref{fig:setup}: 
\begin{itemize}
\item[(i)~~] Centralized $B_i(X_1,\ldots,X_n)$, $i=1,\ldots,n$.
\item[(ii)~] Adaptive $B_i(X_i,B_1,\ldots,B_{i-1})$, $i=2,\ldots,n$.
\item[(iii)] Distributed $B_i(X_i)$, $i=1,\ldots,n$.
\end{itemize}

The performance of an estimator ${\theta}_n \triangleq {\theta}_n(B^n)$ in any of these cases is measured according to the quadratic risk:
\begin{equation}
\label{eq:error_def}
R_n \triangleq \mathbb E\left({\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X_1,\ldots,X_n$ and, whenever available, a prior distribution $\pi(\theta)$ over $\Theta$.  
%It is well known that minmax estimation error 
%\begin{equation}
%R_n = \min \sup_{\theta \in \Theta}  \mathbb E\left[ \left({\theta}_n - \theta \right)^2 | \theta \right],
%\end{equation}
%can be obtained from $R_n$ by assuming a particular least favorable prior distribution on the parameter space $\Theta$ \cite{casella1981}. 
%
The main problems we consider in this paper are the minimal value of \eqref{eq:error_def}, as a function of $n$ and $f(x)$, under different choices of the encoding functions in cases (i), (ii), and (iii). \par

We give particular attention to the ARE of estimators with respect to an asymptotically normal efficient estimator that is not subject to the bit constraint. Specifically, let $\{a_n,\,n\in \mathbb N\}$ be a sequence such that 
\[
\sqrt{a_n}\left({\theta}_n - \theta\right) \overset{d}{\longrightarrow} \Ncal(0, \sigma^2).
\]
Then the ARE of ${\theta}_n$ with respect to an unconstrained efficient estimator for $\theta$ is defined as \cite[Def. 6.6.6]{lehmann2006theory}
\[
\ARE({\theta}_n) \triangleq
\lim_{n\rightarrow \infty} \frac{a_n}{n}. 
\]
Note that in the special case where there exists $V \in \mathbb R$ such that
\[
a_n \mathbb E \left({\theta}_n - \theta \right)^2 = V + o(1),
\]
with $o(1)\to 0$ as $n\to \infty$, then the ARE of ${\theta}_n$ is finite and equals $\sigma^2/V$. \par

In addition to the notation above, we also denote by $F(x)$ the cumulative distribution function of $X_i$, and define
\begin{equation} \label{eq:eta_def}
\eta(x) \triangleq \frac{f^2(x)}{F(x)(1-F(x))} =  \frac{f(x)f(-x)}{F(x)F(-x)}, 
\end{equation}
where the last equality is due to symmetry of $f(x)$. We note that 
\begin{equation}
\label{eq:eta_h}
\eta(x) = h(x)h(-x) = f(x) \left( h(x) + h(-x) \right), 
\end{equation}
where 
\[
h(x) \triangleq \frac{f(x)}{1-F(x)} = \frac{f(x)}{F(-x)}
\]
is the \emph{hazard} function (a.k.a. \emph{failure rate} or \emph{force of mortality}), which is a monotone increasing function since $f(x)$ is log-concave \cite{bagnoli2005log}. 
%
For $f(x)$ the normal density, it is shown in \cite{Samford1953} and \cite{hammersley1950estimating} that $\eta(x)$ is a strictly decreasing function of $|x|$, as illustrated in Fig.~\ref{fig:eta}. 
%
In this paper we only consider the normal distribution and other log-concave symmetric distributions for which this property of $\eta(x)$ holds. Specifically, we require the following:
\begin{assump} \label{assump:failure_rate}
 $\eta(x)$ has a unique maximum (at the origin) and is a non-increasing function of $|x|$. 
\end{assump}
%That is, Assumption~\ref{assump:failure_rate} says that $\eta'(x) < 0 $ for all $x>0$. 
Under this assumption we have, 
\[
4f^2(x) \leq \eta(x) \leq \eta(0),
\] 
%
where $\eta(0) = 4 f^2(0)$ is the asymptotic variance of the sample median. Combined with log-concavity of $f(x)$, Assumption~\ref{assump:failure_rate} implies that $\eta(x)$ vanishes as $|x|\rightarrow \infty$. \par
%
Assumption~\ref{assump:failure_rate} is satisfied, for example, by the generalized normal distributions with a shape parameter between $1$ and $2$ (including normal and Laplace distributions). Symmetric log-concave distributions that do not satisfy Assumption~\ref{assump:failure_rate} include the uniform distribution and the generalized normal distribution with shape parameter greater than $2$.\par

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale = 0.6]
\begin{axis}[
width=8cm, height=6cm,
xmin = -3, xmax=3, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 0.9,
samples=10, 
xlabel= $x$,
xtick={-2,-1,0,1,2},
xticklabels={-2,-1,0,1,2},
ytick={0,0.3989423,0.6366198},
yticklabels={0,$\frac{1}{\sqrt{2 \pi}}$,$2/\pi$},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, solid, smooth] plot table [x = x, y = y, col sep=comma] {./Figs/eta.csv};
\addlegendentry{$\eta(x)$};
\addplot[domain = -5:5, samples = 50, color = red, solid, smooth]  {exp(-x^2/2) / sqrt(2*3.14159)};
\addlegendentry{$\phi(x)$};
\addplot[domain = -5:5, samples = 50, color = black, solid, dashed]  {4*exp(-x^2) / (2*3.14159)};
\addlegendentry{$4\phi^2(x)$};

\end{axis}
\end{tikzpicture}
%
\begin{tikzpicture}[scale = 0.6]
\begin{axis}[
width=8cm, height=6cm,
xmin = -4, xmax=4, 
restrict y to domain = -10:0,
ymin = -10,
samples=10, 
xlabel= $x$,
xtick={-3,-2,-1,0,1,2,3},
xticklabels={-3,-2,-1,0,1,2,3},
ytick={0,-0.45158,-0.919},
yticklabels={0,,},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, solid, smooth] plot table [x = x, y = logy, col sep=comma] {./Figs/eta.csv};
\addlegendentry{$\log \eta(x)$};
\addplot[domain = -5:5, samples = 30, color = red, solid, smooth]  {-(x)^2/2 -0.9189};
\addlegendentry{$\log \phi(x)$};
\end{axis}
\end{tikzpicture}
\caption{
The function $\eta(x) = f^2(x) / F(x)F(-x)$ for $f(x) = \phi(x)$ the standard normal density.
\label{fig:eta}
}
\end{center}
\end{figure}

\section{Consistent Estimation and Off-the-shelf Bounds \label{sec:preliminary}}
%
In this section we adapt known results to derive lower and upper bounds on the relative efficiency under setting (iii). These bounds establish the following facts:
\begin{itemize}
\item[I.] A consistent estimator with an asymptotically normal distribution always exists in setting (iii), and hence in setting (ii). 
\item[II.] For the normal distribution, the ARE in setting (iii) is at most $3/4$. Namely, under setting (iii), all estimators are strictly inferior compared to the sample mean. 
\end{itemize}

\subsection{Consistent Estimation}
Fix $\theta_0 \in \mathbb R$ and define the $i$th message by 
\[
B_i = \mathbf 1_{X_i>\theta_0}, 
\]
where $\mathbf 1_A$ is the indicator of the event $A$. We have
\[
P_n \triangleq\frac{1}{n} \sum_{i=1}^n B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),  
\]
so that 
\begin{equation}
\label{eq:estimator_naive}
{\theta}_n = \theta_0 + F^{-1}\left( P_n \right)
\end{equation}
is a consistent estimator for $\theta$ in the distributed setting of Figure~\ref{fig:setup}-(iii), where we note that $F(x)$ is invertible over the support of $f(x)$ which is a connected set due to log-concavity. Furthermore, the variance of $P_n$ is $F(\theta-\theta_0)\left(1-F(\theta-\theta_0)\right)$, and hence the delta method implies that ${\theta}_n$ is asymptotically normal with variance 
\begin{equation} 
 \frac{F(\theta-\theta_0)\left(1-F(\theta-\theta_0)\right)}{f^2(\theta-\theta_0)} = \frac{1}{
\eta(\theta-\theta_0)}.
\end{equation}
In particular, the ARE of ${\theta}_n$ equals $\eta(\theta - \theta_0)\sigma^2$. In other words, for a prescribed accuracy, ${\theta}_n$ of  \eqref{eq:estimator_naive} estimates $\theta$ with sample size that is $\eta(\theta- \theta_0)  \sigma^2$ times the samples size required for the sample mean. \par
%
%
Assumption~\ref{assump:failure_rate} implies that for  all $n$ large enough the ARE of ${\theta}_n$ of \eqref{eq:estimator_naive} is never greater than $\eta(0)$, which is the ARE of the sample median. This ARE is attained only when $\theta_0 = \theta$, whereas $\theta$ is apriori unknown. Since $\eta(x)$ vanishes as $|x|\rightarrow \infty$, the ARE of ${\theta}_n$ may be very small when $\theta$ is away from $\theta_0$. As an example, when $f(x)$ is a normal density, the ARE of ${\theta}_n$ is less than $\approx 0.14$ when $\theta_0$ is $2$ standard deviations from $\theta$. Therefore, the estimator ${\theta}_n$ has little practical value unless the radius of $\Theta$ is small compared to the standard deviation. \par 
It is suggested that we can attain lower variance in estimation by updating the threshold value $\theta_0$ in \eqref{eq:estimator_naive} after observing one or a batch of the single bit messages. Schemes with such adaptation fall within the adaptive setting of Figure~\ref{fig:setup}-(ii) which we consider in Section~\ref{sec:sequential}. \par

\subsection{Multiterminal Source Coding \label{sec:ceo}}
The CEO setting considers the estimation of a sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$ is available at $n$ terminals. At each terminal $i$, an encoder observes the $k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $r_i k$ bits to a central estimator  \cite{berger1996ceo}. The central estimator produces an estimates ${\theta}_1,\ldots,{\theta}_k$ with the goal of minimizing the quadratic risk:
\[
R_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - {\theta}_j \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $r_1=\ldots=r_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi(d\theta)$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the distributed encoding setting is given by the minimal MSE in the CEO setting as $k \to \infty$. Note that the difference between the CEO setting and ours lays in the privilege of each of the encoders to describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realizations, rather than a single realization using a single bit in ours. 
 \par
When the prior on $\theta$ and the noise corrupting it at each location are Gaussian, the optimal encoding scheme and its asymptotic risk as $k$ goes to infinity were fully characterized in \cite{prabhakaran2004rate}. Furthermore, the work of \cite{chen2004upper} provided an expression for quadratic risk attained in the CEO setting under Gaussian priors. Adapting to our setting, this expression provides the following proposition:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\Theta = \mathbb R$ and $\pi(\theta) = \Ncal(0,\sigma_\theta^2)$ where $\sigma_\theta^2 \in \mathbb R$ is arbitrary. Then any estimator ${\theta}_n$ of $\theta$ in the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n \ex{\left( \theta - \theta_n \right)^2} \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
\end{equation}
where the expectation is with respect to $\theta$ and $X^n$.
\end{prop}

\begin{proof}
See Appendix~\ref{app:proof:prop_ceo_lower_bound}.
\end{proof}

From the formulation of the CEO problem, it follows that the difference between the MSE lower bound \eqref{eq:ceo_bound} and the actual MSE in the distributed setting (case (iii)) is attributed to the ability to perform coding over blocks. Namely, each encoder in the CEO setting may encode an arbitrary number of $k$ independent realizations of $\theta$ using $k$ bits, versus only one realization with one bit in ours. In other words, it is the ability to exploit the geometry of a high-dimensional product probability space that distinguishes between the CEO problem with one bit per encoder on average and the mean estimation problem from one-bit measurements in the distributed setting.

\section{Adaptive Estimation \label{sec:sequential}}
The first main results of this paper, as described in Theorem~\ref{thm:adpative_lower_bound} below, states that the ARE of any adaptive estimator cannot be larger than $\eta(0)\sigma^2$, which is the ARE of the median of the sample $X_1,\ldots,X_n$. We also provide a particular adaptive estimation scheme that attains this maximal efficiency. %Finally, in Theorem~\ref{thm:opt_one_step}, we provide an adaptive estimation scheme that is one-step optimal in the sense that at each step $i$, the chosen message $B_i$  minimizes the MSE given $X_i$ and the previous $i-1$ messages. %Numerical simulations for the case of $f(x)$ the normal density function shows that the ARE of the estimator described by this scheme is $\eta(0)\sigma^2$.

\subsection{Limited efficiency in the adaptive setting}
Our first result asserts that the ARE of any adaptive encoding and estimation scheme is bounded from above by $\eta(0)\sigma^2$. %This result follows from the following theorem:
\begin{thm}[maximal relative effeciency] \label{thm:adpative_lower_bound}
Let ${\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Figure~\ref{fig:setup}-(ii). Assumes that $\pi(\theta)$ converges to zero at the endpoints of the interval $\Theta$. Then
\[
n\mathbb E\left[ (\theta-{\theta}_n)^2 \right] \geq   \frac{n}{ 4f^2(0) n + I_0},
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log \pi (\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

>>>>

% -*- Mode: latex -*- %

\section{Introduction}
\label{sec:Intro}

We consider estimation of parameters from data collected by multiple units
under communication constraints between the units.  Such scenarios arise in
sensor arrays, where sensor motes collect information, which they transmit
to a central estimation unit~\cite{LesserOrTa03,LiWoHuSa02}. More generally,
communication is substantially more expensive than computation in modern
computing infrastructure~\cite{FullerMi11}.  It is thus of interest to
understand the extent to which communication constraints induce fundamental
accuracy and efficiency limits in parametric estimation problems.
%%
\begin{figure*}
  \include{Figs/encoding-settings}
  \caption{\label{fig:setup} Three encoding settings: (i) Centralized -- an
    encoder sends $n$ bits after observing $n$ samples. (ii) Adaptive
    (sequential) -- the $i$th encoder sends the bit $B_i$ depending on its
    private sample $X_i$ and previous bits $B_1,\ldots,B_{i-1}$. (iii)
    Distributed -- each encoder send the bit $B_i$ based on its private
    sample $X_i$ only.}
\end{figure*}
%%

We answer this question in a sylized version of this problem: the estimation
of the mean $\theta$ of a symmetric log-concave distribution under the
constraint that only a single bit can be communicated about each observation
from this distribution.

%% JCD: I am here. %%

As it turns out, the ability to share information
before committing on each one-bit message dramatically affects the
performance in estimating $\theta$. We, therefore, distinguish among the
three settings illustrated in Figure~\ref{fig:setup}
\begin{itemize}
 \item[(i)\,\,] \emph{Centralized} encoding: all $n$ encoders confer and produce a single message consists of $n$ bits.
 \item[(ii)\,] \emph{Adaptive} or \emph{sequential} encoding: The $n$th encoder observes the $n$th sample and the $n-1$ previous bits.
 \item[(iii)] \emph{Distributed} encoding: The $n$th message is only a function of the $n$th sample.
 \end{itemize}
Evidently, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). Below are three application examples for each of the settings (i)-(iii) above, respectively:
\begin{itemize}
\item {\bf Signal acquisition:} A quantity is measured $n$ times at different instances. The results are averaged in order to reduce measurement noise and the averaged result is then stored or communicated using $n$ bits. 
\item {\bf Analog-to-digital conversion:} A sigma-delta modulator (SDM) converts an analog signal into a sequence of bits by sampling the signal at a very high rate and then using one-bit threshold detector combined with a feedback loop to update an accumulated error state \cite{1092194}. Therefore, the expected error in tracking an analog signal using an SDM falls under our setting (ii) when we assume that the signal at the input to the modulator is a constant (direct current) corrupted by, say, thermal noise \cite{53738}. Since the sampling rates in SDM are usually many times more than the bandwidth of its input, analyzing SDM under a constant input provides meaningful lower bound even for non-constant signals.
\item {\bf Privacy:} A business entity is interested in estimating the average income of its clients. In order to keep this information as confidential as possible, each client independently provides an answer to a yes/no question related to its income.
\end{itemize}
%
We measure the performance in estimating $\theta$ by the quadratic risk. For an estimator with finite quadratic risk $R_n$, we are interested in particular in the limit
\begin{equation}
\label{eq:ARE_def}
\lim_{n\to \infty} \frac{\sigma^2/n}{R_n},
\end{equation}
describing the \emph{asymptotic relative efficiency} (ARE) of this estimator compared to an estimator whose variance decreases as $\sigma^2/n+o(n^{-1})$. Estimators of this form include the empirical mean of the samples, and, under some conditions, the optimal Bayes estimator. \par
%
In addition to the examples above, the excess risk in estimating a fixed parameter due to a one-bit per sample constraint is useful in bounding from below the excess risk in estimating signals from thier noisy one-bit quanitzed noisy measurements. Specifically, the excess risk or ARE we derive serve as the most optimistic estimate for the risk in estimating under one-bit per measurement constraint signals changing over time or space. Such estimation settings are considered in \cite{baraniuk2017exponential, jacques2013robust, plan2013one, li2017channel, choi2016near}. \par

@@
